Analysing Constraint Grammars with a SAT-solver

### 2015-10-25 paper/lrec/eval_futurework.tex
remove more words

We tested three grammars to find rules that cannot apply. 
The smallest grammar was
Dutch\footnote{\url{https://svn.code.sf.net/p/apertium/svn/languages/apertium-nld/apertium-nld.nld.rlx}},
with 59 rules; second was
Spanish\footnote{\url{https://svn.code.sf.net/p/apertium/svn/languages/apertium-spa/apertium-spa.spa.rlx}},
with 279 rules, and the largest was
Finnish\footnote{\url{https://github.com/flammie/omorfi/blob/master/src/vislcg3/omorfi.cg3}},
with 1185 rules. We left out \textsc{add}, \textsc{map} and other rule
types introduced in CG-3, and only tested the traditional
\textsc{remove} and \textsc{select} rules.
The results are shown in table~\ref{table:res}.

@@

\paragraph{Correctness} 
The test revealed problems in all grammars, and for the smaller
grammars, we were able to verify manually that the detected rules were
true negatives. However, for the Finnish grammar, we did not have time
to investigate all of them; in addition, we suspect that some of
them are false negatives.
We did not check for false positives in any of the grammars.

@@

For the Finnish grammar, we could verify a number of tagset errors. 
The grammar was originally written in 1995, and converted to the
Omorfi tagset by \cite{pirinen2015}. For example, the Omorfi tagset
marks a possessive suffix with \texttt{Px}, e.g. \texttt{PxPl2} for second person plural, but the
grammar used the old convention \texttt{PL2} in the rules.
As a result, all rules with possessive suffixes were detected as
impossible.

Because of performance issues, we ignored the readings from the
lexicon and used only tag combinations that were named in the grammar.
For example, the rule \texttt{REMOVE (verb sg) IF (-1 det)}
gives us ``verb sg'' and ``det'', as opposed to fully specified
readings in the lexicon.
The shortcut will work most of the time, but it is possible to
miss some cases: e.g. \texttt{SELECT (verb sg) IF (0 p3)} 
requires a full ``verb sg p3'' reading, but our method only gives
``verb sg'' and ``p3'' separately. 
Some of the false negatives seem to come from this shortcut, or our naive
implementation of rules with \verb!*! in their context.
Due to these reasons, we cannot give conclusions about the Finnish
grammar; we include the results here only to have some preliminary
data on scaling up to larger grammars.
%For a condition situated $n$ or more words to the left or right, we only create a word exactly $n$ words away. This is something to be fixed in the future.

\paragraph{Running time} The speed of the check is an obvious issue,
especially for the Finnish grammar. 
As mentioned before, we took a shortcut and used 2000 readings from
the grammar.
With the full set of 6000 readings from the lexicon, the speed
was even worse: checking the last rule took 30 minutes, as opposed to
3 minutes with 2000 readings. 
%We never finished running the whole rule set with 6000 readings, but it would have taken potentially days.
However, the code is currently just a proof of concept, and we plan to
make an effort in optimisation.

@@

We will work on the performance, and fix any errors we found in the
evaluation, such as the issue with non-exact position.
Another obvious improvement is to restrict possible ambiguities
in the symbolic words.
Currently we have only hard-coded some
very basic assumptions, such as ``word cannot be ambiguous between a
boundary and a real word'', but we want to extract these restrictions
automatically.
For instance, the Spanish lexicon contains the entries
\texttt{casa:casa<n><sg>}  and  \texttt{casa:casar<v><sg><p3>}, thus
we know that a word can be ambiguous between a noun and a verb.
By adding these restrictions to the symbolic words, we hope to
restrict the search space, and make the SAT problem easier to solve.

%We will also improve usability, by finding out which rules contrad
After this, we want to handle the full expressivity of CG-3, with \textsc{map}, \textsc{add} and \textsc{substitute} rules.
At that stage, it would be interesting to test our program with actual grammar writers,
in comparison with a corpus-based method or machine learning.
Finally, if the method proves feasible for Constraint Grammars, we want
to try applying it to other, more complex grammar formalisms.

>>>>

We tested three grammars to find rules that cannot apply: 
Dutch\footnote{\url{https://svn.code.sf.net/p/apertium/svn/languages/apertium-nld/apertium-nld.nld.rlx}},
with 59 rules; 
Spanish\footnote{\url{https://svn.code.sf.net/p/apertium/svn/languages/apertium-spa/apertium-spa.spa.rlx}},
with 279 rules; and 
Finnish\footnote{\url{https://github.com/flammie/omorfi/blob/master/src/vislcg3/omorfi.cg3}},
with 1185 rules. We left out \textsc{add}, \textsc{map} and other rule
types introduced in CG-3, and only tested \textsc{remove} and \textsc{select} rules.
The results are shown in table~\ref{table:res}.

@@

\paragraph{Correctness} 
The test revealed problems in all grammars, and for the smaller
grammars, we were able to verify manually that the detected rules were
true negatives.
%The results for the Finnish grammar are inconclusive, 
% For the Finnish grammar, we did not have time to
% investigate all of them, but among those we could, we found both true and false negatives.
We did not check for false positives in any of the grammars.

@@

For the Finnish grammar, we could verify a number of tagset errors. 
The grammar was originally written in 1995, and converted to the
Omorfi tagset by \cite{pirinen2015}. For example, the Omorfi tagset
marks a possessive suffix with \texttt{Px}, e.g. \texttt{PxPl2} for second person plural, but the
grammar uses the old convention \texttt{PL2} in the rules.
As a result, all rules with possessive suffixes were detected as
impossible.

Because of performance issues, we ignored the fully specified readings from the
lexicon and used only tag combinations from the grammar:
for example, the rule \texttt{REMOVE (verb sg) IF (-1 det)}
gives us ``verb sg'' and ``det''.
The shortcut works most of the time, but it is possible to
miss some cases: e.g. \texttt{SELECT (verb sg) IF (0 p3)} 
assumes ``verb sg p3'', but our method only gives
``verb sg'' and ``p3'' separately. 
Some of the detected rules seem to be false negatives caused by this shortcut. 
Another suspect is our naive implementation of rules with \verb!*! in
their context.

Due to these reasons, the results about the Finnish grammar are
inconclusive. We only include them to show some preliminary data on larger grammars.
%For a condition situated $n$ or more words to the left or right, we only create a word exactly $n$ words away. This is something to be fixed in the future.

\paragraph{Performance} The speed of the check is an obvious issue,
especially for the Finnish grammar. 
As mentioned before, we used 2000 readings extracted from the grammar.
With the full set of 6000 readings from the lexicon, the speed
was even worse: checking the last rule took 30 minutes, as opposed to
3 minutes with 2000 readings. 
%We never finished running the whole rule set with 6000 readings, but
%it would have taken potentially days.
We plan to make an effort in optimising the code; currently it is just
a naive proof-of-concept implementation.

@@

We will work on the performance, and fix the errors we found in the
evaluation, such as the issue with non-exact position.
We also plan to restrict possible ambiguities in the symbolic words. 
Currently we have only hard-coded some
very basic assumptions, such as ``word cannot be ambiguous between a
boundary and a real word'', but we want to extract these constraints
automatically from a lexicon.
% For instance, the Spanish lexicon contains the entries
% \texttt{casa:casa<n><sg>}  and  \texttt{casa:casar<v><sg><p3>}, thus
% we know that a word can be ambiguous between a noun and a verb.
By adding these restrictions to the symbolic words, we hope to
restrict the search space, and make the SAT problem easier to solve.

%We will also improve usability, by finding out which rules contrad
For longer-term goals, we want to handle the full expressivity of CG-3,
with \textsc{map}, \textsc{add} and \textsc{substitute} rules, and
dependency structure.
As soon as the tools are mature enough, we want to
evaluate them with actual grammar writers,
in comparison with a corpus-based method or machine learning.
Finally, if the method proves feasible for CG, we want
to try applying it to other grammar formalisms.

### 2015-10-25 paper/lrec/problem_desc.tex
remove more words

CGs are valuable resources for rule-based NLP, especially for lesser resourced languages. They are robust and can be written without large corpora--only morphological analysis is needed. The formalism is lightweight and language-independent, and resources can be shared between related languages \cite{bick2006spanish,lene_trond_linda2010}.
%Mature CGs contain some thousands of rules, but even small CGs are shown to be effective \cite{lene_trond2011}.

@@

\begin{itemize}
\item Corpus-based methods in manual grammar development \cite{voutilainen2004}
\item Optimising hand-written grammars \cite{bick2013tuning}
\item Encoding CG in logic \cite{lager98,lager_nivre01,listenmaa_claessen2015}
\end{itemize}

In addition, there is a large body of research on automatically inducing rules, e.g. \cite{inducing_cg1996}, \cite{lindberg_eineborg98ilp}
%, \cite{lager01transformation}, \cite{asfrent14} 
. However, since our work is aimed to aid the process of hand-crafting rules, we omit those works from our discussion.

@@

\cite{bick2013tuning} modifies the grammar automatically, by trying out different orderings of the rules, and changing their contexts to stricter or less strict.
Bick reports error reduction of 7--15\% compared to the original grammars.
As a downside, the grammar writer will likely not know why exactly is the tuned grammar performing better.
% At a certain point, the grammar gets so big that it is hard to keep track of all the rules and their interactions. \cite{bick2013tuning} tries out combinations of moving rules in different sections or removing them in total, and in parallel, making their contexts stricter or less strict. 
% This is a valuable tool, especially for grammars that are so big that it's hard to keep track of. Program can try all combinations whereas trying to make sense out of a huge set of rules would be hard for humans.

>>>>

CGs are valuable resources for rule-based NLP, especially for lesser
resourced languages. They are robust and can be written without large
corpora---only morphological analysis is needed. The formalism is
lightweight and language-independent, and resources can be shared
between related languages \cite{bick2006spanish}.
%,lene_trond_linda2010}.
%Mature CGs contain some thousands of rules, but even small CGs are shown to be effective \cite{lene_trond2011}.

@@

\begin{itemize}
\item Corpus-based methods in manual grammar development \cite{voutilainen2004}
\item Optimising hand-written CGs~\cite{bick2013tuning}
\item Encoding CG in logic \cite{lager98,lager_nivre01,listenmaa_claessen2015}
\end{itemize}

In addition, there is a large body of research on automatically inducing rules, e.g. \cite{inducing_cg1996,lindberg_eineborg98ilp}.
%, \cite{lager01transformation}, \cite{asfrent14} 
However, since our work is aimed to aid the process of hand-crafting rules, we omit those works from our discussion.

@@

\cite{bick2013tuning} modifies the grammar automatically, by trying
out different rule orders and altering the contexts of the rules.
Bick reports error reduction of 7--15\% compared to the original grammars.
As a downside, the grammar writer will likely not know why exactly does the tuned grammar perform better.
% At a certain point, the grammar gets so big that it is hard to keep track of all the rules and their interactions. \cite{bick2013tuning} tries out combinations of moving rules in different sections or removing them in total, and in parallel, making their contexts stricter or less strict. 
% This is a valuable tool, especially for grammars that are so big that it's hard to keep track of. Program can try all combinations whereas trying to make sense out of a huge set of rules would be hard for humans.
